---
title: "CFPB_Project"
author: "Ashwini"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
# Loading raw cfpb_data
cfpb_equifax_data <- read.csv("D:\\US Master's\\Fall 2023\\Ideas and Principles of Data Science\\Lecture_Dataset\\complaints-2023-09-16_14_53.csv")
dim(cfpb_equifax_data)
```

### Data cleaning

```{r}
library(DataExplorer)
plot_intro(cfpb_equifax_data)
```

```{r}
head(cfpb_equifax_data)
```
### Columns in CFPB_data
```{r}
colnames(cfpb_equifax_data)
```
# Data description of columns
- consumer_complaint_narrative: Text written by the cutsomer pertaining to the complaint
- product : The high level financial product to which the complaint relates
- sub_product :	A specific product within the high level product category
- issue:	The type of complaint being raised
- sub_issue:	The low level type of issue submitted
- consumer_consent_provided:	Binary of whether the consumer provided consent to be contacted by the company
- timely_response	: Whether the company responded to the complaint within 15 days
- submitted_via:	The channel by which the complaint was lodged
- company_public_response:	How the company publically responded to the complaint
- company:	The company’s name against which the complaint was lodged
- company_response_to_consumer:	How the company responded to the customer
- consumer_disputed	: Did the consumer dispute the companies response/proposed resolution

# renaming the columns to snake_case variable
```{r}
library(dplyr)
new <- c('date_received','product','sub_product','issue','sub_issue','consumer_complaint_narrative','company_public_response','company','state','zip_code','tags','consumer_consent_provided','submitted_via','date_sent_to_company','company_response_to_consumer','timely_response','consumer_disputed','complaint_id')

old <- c('Date.received','Product','Sub.product','Issue','Sub.issue','Consumer.complaint.narrative','Company.public.response','Company','State','ZIP.code','Tags','Consumer.consent.provided.','Submitted.via','Date.sent.to.company','Company.response.to.consumer','Timely.response.','Consumer.disputed.','Complaint.ID')

#rename old names with new names
cfpb_equifax_data <- cfpb_equifax_data %>% rename_with(~ new, all_of(old))

head(cfpb_equifax_data)
                                                
```
### Count of unique values in columns
```{r}
rapply(cfpb_equifax_data, function(x) length(table(x)))

```
### Replace of None and N/A as NaN
```{r}
cfpb_equifax_data[cfpb_equifax_data == "None" | cfpb_equifax_data == "N/A" | cfpb_equifax_data == ""] <- NA

head(cfpb_equifax_data)
```
### Plot representing missing values in columns

```{r}
plot_intro(cfpb_equifax_data)
```

```{r}
plot_missing(cfpb_equifax_data)
```

### Unique values in columns marked in purple in above plot
```{r}
print("timely response-")
print(unique(cfpb_equifax_data$timely_response))
print("consumer disputed category-")
print(unique(cfpb_equifax_data$consumer_disputed))
print("--------------------------------------------")
print("tags-")
print(unique(cfpb_equifax_data$tags))
print("--------------------------------------------")
print("company public response-")
print(unique(cfpb_equifax_data$company_public_response))
```
### Replacing the NA by Unknown in columns 'Consumer_disputed_category','tags' and 'company_public_response'
```{r}
# Fill NaN by "Unknown" in column consumer_disputed
library("dplyr")
cfpb_equifax_data <- cfpb_equifax_data %>%
  mutate(consumer_disputed = coalesce(consumer_disputed, "Unknown"),
         tags = coalesce(tags, "Unknown"),
         consumer_consent_provided = coalesce(consumer_consent_provided, "Other"),
         sub_product = coalesce(sub_product,"Unknown"))


# Clean zip_code col using package zipcodeR
library(zipcodeR)
zipcodes_zcta <- is_zcta(cfpb_equifax_data$zip_code)
cfpb_equifax_data <- subset(cfpb_equifax_data, zipcodes_zcta == TRUE)

```

```{r}
sum(is.na(cfpb_equifax_data$state))
```

# Check whether the changes are reflected
```{r}
print(unique(cfpb_equifax_data$consumer_disputed))
```
### Deleting the cols
company - Equifax Inc.
complaint id: id of the complaint
sub_issue: beneficial in text analysis
company_public_response: 99.97% vol of complaints are NaN

```{r}
drops <- c("company","complaint_id","company_public_response","sub_issue")
cfpb_equifax_data <-cfpb_equifax_data[ , !(names(cfpb_equifax_data) %in% drops)]
head(cfpb_equifax_data)
```

# Remove rows which have X in the column "zipcode"


```{r}
#zip_df <- geocode_zip(c(cfpb_equifax_data$zip_code))
#colnames(zip_df)[colnames(zip_df) == 'zipcode'] <- 'zip_code'
#cfpb_equifax_data <- left_join(cfpb_equifax_data,zip_df,by='zip_code')

```

We are still left with two more variables to clean, we shall deal with it in the later course.

### Converting the date column to date type
```{r}
library(lubridate)
d1 <- mdy(cfpb_equifax_data$date_received)
d2 <- mdy(cfpb_equifax_data$date_sent_to_company)
cfpb_equifax_data$date_received <- d1
cfpb_equifax_data$date_sent_to_company <- d2

# Subset of data from 2018 to Sep-2023
cfpb_equifax_data <- subset(cfpb_equifax_data, date_received >= as.Date('2018-01-01'))
cfpb_equifax_data$year <- format(as.Date(cfpb_equifax_data$date_received, format="%Y-%m-%d"),"%Y")
head(cfpb_equifax_data)
```

### Sort the values based on date_received in ascending order
```{r}
cfpb_equifax_data <- cfpb_equifax_data[order(as.Date(cfpb_equifax_data$date_received, format="%Y-%m-%d")),]
```

# Difference in days from complaint_recieved to complaint_sent_to_company

```{r}
cfpb_equifax_data$days_diff = difftime(cfpb_equifax_data$date_sent_to_company,cfpb_equifax_data$date_received,units='days')
cfpb_equifax_data <- transform(cfpb_equifax_data, days_diff = as.numeric(days_diff))
```

```{r}
cfpb_equifax_data[cfpb_equifax_data$days_diff==288,]
```


Further looking at the days where the complaints being received across all products.

```{r}
library(ggplot2)
cfpb_equifax_data$day <- weekdays(as.Date(cfpb_equifax_data$date_received))

# Add a new column indicating whether the day is a weekday or weekend
ggplot(data = cfpb_equifax_data, aes(y = factor(day))) +
  geom_bar(fill = "steelblue") +
  labs(title = "Bar Plot of days", x = "Count", y = "day")

```
We can observe here that Wednesdays ,Tuesdays and thursdays have lot of complaints being received compared to the other days. Reason could be people are no more in their relaxing mood and want to follow-up on the concerns they have with CFPB which aren't being addressed by EQUIFAX INC.

# Volume of complaints submitted via various medium across years
```{r}
table(cfpb_equifax_data$submitted_via,cfpb_equifax_data$year)
```


```{r}
# Creating new feature from col consumer_complaint_narrative
cfpb_equifax_data$narrative_count <- nchar(cfpb_equifax_data$consumer_complaint_narrative)
# replace any nan with 0 
cfpb_equifax_data <- cfpb_equifax_data %>% mutate(narrative_count = ifelse(is.na(narrative_count), 0, narrative_count))
head(cfpb_equifax_data)
```

# Creation of an indicator variable for tracking covid-19
```{r}
cfpb_equifax_data <- cfpb_equifax_data %>%
  mutate(covid_indicator = ifelse(date_received >= as.Date("2020-01-01") & date_received <= as.Date("2020-06-30"), 1, 0))
```


We can observe over the years Web is the most preferred medium to submit complaints to CFPB.


```{r}
# Unique products of complaints
unique(cfpb_equifax_data$product)
```
```{r}
# Reducing the no of product categories
library(dplyr)

cfpb_equifax_data <- within(cfpb_equifax_data, {
       product[product == "Credit reporting or other personal consumer reports"] <- "Credit reporting, credit repair services, or other personal consumer reports"
       product[product == "Credit card"] <- "Credit card or prepaid card"
       product[product == "Student loan"] <- "Other loan"
       product[product == "Payday loan, title loan, or personal loan"] <- "Other loan"
       product[product == "Vehicle loan or lease"] <- "Other loan"
})

```

```{r}
prd_cnt <- cfpb_equifax_data %>% count(product,sort=TRUE)
print(prd_cnt)
```

# Checking the distribution of product plot
```{r}
library(dplyr)
prd_cnt <- cfpb_equifax_data %>% count(product,sort=TRUE)
ggplot(prd_cnt, aes(y = product, x = n)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Count of Products complaint for Equifax",
       x = "Product",
       y = "Count") +
  theme_minimal() 
#theme(axis.text.y = element_text(angle = 75, hjust = 1)) 
```

We will be working on the first 3 products as it consumes the maximum volume of complaints.
```{r}
cfpb_equifax_data <- cfpb_equifax_data[cfpb_equifax_data$product %in% c('Credit reporting, credit repair services, or other personal consumer reports', 'Debt collection', 'Credit card or prepaid card'), ]
```

We often see that most of the values are repeated, we shall classify the severity of issues based on the volume and frequency reported over years

```{r}
unique(cfpb_equifax_data$issue)
```
# Creating target variable based on column Issue and classifying the severity 
# issue as Low , Medium and High

```{r}
cfpb_equifax_data <- cfpb_equifax_data %>%
mutate(issue_severity = case_when(issue == "Attempts to collect debt not owed" |
  issue  == "Credit monitoring or identity theft protection services" |
  issue  == "Improper use of your report" |
  issue == "Incorrect information on your report" |
  issue == "Problem with a credit reporting company's investigation into an existing problem" |
  issue == "Problem with fraud alerts or security freezes" |
  issue == "Unable to get your credit report or credit score" ~ 'high',
  issue == "False statements or representation" |
  issue == "Fraud or scam" |
  issue == "Getting a credit card" |
  issue == "Identity theft protection or other monitoring services" |
  issue == "Problem with a company's investigation into an existing issue" |
  issue == "Took or threatened to take negative or legal action" |
  issue == "Written notification about debt" ~ 'medium',
  issue == "Advertising" |
  issue == "Advertising and marketing, including promotional offers" |
  issue == "Closing an account" |
  issue == "Closing your account" |
  issue == "Communication tactics" |
  issue == "Confusing or misleading advertising or marketing" |
  issue == "Confusing or missing disclosures" |
  issue == "Excessive fees" |
  issue == "Fees or interest" |
  issue == "Getting a loan" |
  issue == "Getting a loan or lease" |
  issue == "Other features, terms, or problems" |
  issue == "Problem getting a card or closing an account" |
  issue == "Problem when making payments" |
  issue == "Problem with a company's investigation into an existing problem" |
  issue == "Problem with a purchase or transfer" |
  issue == "Problem with a purchase shown on your statement" |
  issue == "Problem with customer service" |
  issue == "Problem with overdraft" |
  issue == "Struggling to pay your bill" |
  issue == "Threatened to contact someone or share information improperly" |
  issue == "Trouble using the card" |
  issue == "Trouble using your card" |
  issue == "Unexpected or other fees" ~ 'low'))

```


# Drop of target column issue

```{r}
# Drop cols consumer
drops <- c("consumer_complaint_narrative","year","date_received","date_sent_to_company")
cfpb_equifax_data <-cfpb_equifax_data[ , !(names(cfpb_equifax_data) %in% drops)]
```

```{r}
cfpb_equifax_data <- na.omit(cfpb_equifax_data)
```

# Creation of new dataframes based on top 3 volume products
```{r}
# Filter out data for product "credit card or prepaid card"
card_df <- cfpb_equifax_data[cfpb_equifax_data$product == "Credit card or prepaid card",]
credit_df <- cfpb_equifax_data[cfpb_equifax_data$product == "Credit reporting, credit repair services, or other personal consumer reports",]
debt_df <- cfpb_equifax_data[cfpb_equifax_data$product == "Debt collection",]
```


```{r}
head(card_df)
```

# Distribution of target variable issue_severity across 3 dataframes
```{r}
library(ggplot2)
library("ggpubr")

a <- ggplot(card_df,aes(x=reorder(issue_severity,issue_severity,function(x)-length(x)))) +
  geom_bar(fill='darkgreen') + labs(x='severity of issue across credit or prepaid cards')

b <- ggplot(credit_df,aes(x=reorder(issue_severity,issue_severity,function(x)-length(x)))) +
  geom_bar(fill='grey') + labs(x='severity of issue across credit reporting')

c <- ggplot(debt_df,aes(x=reorder(issue_severity,issue_severity,function(x)-length(x)))) +
  geom_bar(fill='blue') + labs(x='severity of issue across debt collection')

figure <- ggarrange(a,b,c,labels=c("A","B","C"),ncol=2,nrow=2)
figure
```

We observe a high bias among issue severity in Product "Credit reporting"
```{r}
credit_df %>% count(issue,sort=TRUE)
```
```{r}
credit_df <- subset(credit_df, !(issue %in% c("Incorrect information on your report", "Problem with a credit reporting company's investigation into an existing problem","Improper use of your report")))
```

# Rechecking the distribution of target variable
```{r}
library(ggplot2)
library("ggpubr")

a <- ggplot(card_df,aes(x=reorder(issue_severity,issue_severity,function(x)-length(x)))) +
  geom_bar(fill='darkgreen') + labs(x='severity of issue across credit or prepaid cards')

b <- ggplot(credit_df,aes(x=reorder(issue_severity,issue_severity,function(x)-length(x)))) +
  geom_bar(fill='grey') + labs(x='severity of issue across credit reporting')

c <- ggplot(debt_df,aes(x=reorder(issue_severity,issue_severity,function(x)-length(x)))) +
  geom_bar(fill='blue') + labs(x='severity of issue across debt collection')

figure <- ggarrange(a,b,c,labels=c("A","B","C"),ncol=2,nrow=2)
figure
```

```{r}
library(zipcodeR)
library(leaflet)

zipcode_df <- reverse_zipcode(card_df$zip_code)
zipcode_df <- zipcode_df[,c("zipcode","lat","lng")]

card_df <- left_join(card_df, zipcode_df, by = c("zip_code" = "zipcode"))
credit_df <- left_join(credit_df, zipcode_df, by = c("zip_code" = "zipcode"))
debt_df <- left_join(debt_df,zipcode_df, by = c("zip_code" 
                                                = "zipcode"))

card_df$issue_severity <- factor(card_df$issue_severity, levels = c("low", "medium", "high"), ordered = TRUE)

credit_df$issue_severity <- factor(credit_df$issue_severity, levels = c("low", "medium", "high"), ordered = TRUE)

debt_df$issue_severity <- factor(debt_df$issue_severity, levels = c("low", "medium", "high"), ordered = TRUE)
```


```{r}
# Convert frequencies to colors using RColorBrewer
library(RColorBrewer)
color_palette <- brewer.pal(n = nlevels(card_df$issue_severity),name="Dark2")
colors <- colorFactor(color_palette, levels = levels(card_df$issue_severity), ordered = TRUE)

# Plot on a map using leaflet
leaflet(card_df) %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  addCircleMarkers(
    lng = ~lng,
    lat = ~lat,
    popup = ~paste("zipcode: ", zip_code, "<br>Issue severity: ", issue_severity),
    color = ~colors(issue_severity),
    fillOpacity = 0.8,
    radius = 8
  ) %>%
  addLegend(
    position = "bottomright",
    pal = colors,
    values = ~issue_severity,
    title = "Severity of Issue in Credit card or prepaid card"
  )

```


```{r}
# Convert frequencies to colors using RColorBrewer
library(RColorBrewer)
color_palette <- brewer.pal(n = nlevels(credit_df$issue_severity), name = "Dark2")
colors <- colorFactor(color_palette, levels = levels(credit_df$issue_severity), ordered = TRUE)

# Plot on a map using leaflet
leaflet(credit_df) %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  addCircleMarkers(
    lng = ~lng,
    lat = ~lat,
    popup = ~paste("zipcode: ", zip_code, "<br>Issue severity: ", issue_severity),
    color = ~colors(issue_severity),
    fillOpacity = 0.8,
    radius = 8
  ) %>%
  addLegend(
    position = "bottomright",
    pal = colors,
    values = ~issue_severity,
    title = "Severity of Issue in Credit reporting"
  )
```


```{r}
# Convert frequencies to colors using RColorBrewer
library(RColorBrewer)
color_palette <- brewer.pal(n = nlevels(debt_df$issue_severity), name = "Dark2")
colors <- colorFactor(color_palette, levels = levels(debt_df$issue_severity), ordered = TRUE)

# Plot on a map using leaflet
leaflet(debt_df) %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  addCircleMarkers(
    lng = ~lng,
    lat = ~lat,
    popup = ~paste("zipcode: ", zip_code, "<br>Issue severity: ", issue_severity),
    color = ~colors(issue_severity),
    fillOpacity = 0.8,
    radius = 8
  ) %>%
  addLegend(
    position = "bottomright",
    pal = colors,
    values = ~issue_severity,
    title = "Severity of Issue in Debt collection"
  )
```



```{r}
# relationship among narrative count and severity
par(mfrow=c(1,3))
boxplot(narrative_count~issue_severity,data=card_df)
title("Credit or prepaid card")
boxplot(narrative_count~issue_severity,data=credit_df)
title("Credit reporting")
boxplot(narrative_count~issue_severity,data=debt_df)
title("Debt collection")
```

```{r}
par(mfrow=c(1,3))
x <- card_df[card_df$days_diff>= 5,]
y <- credit_df[credit_df$days_diff>=5,]
z <- debt_df[debt_df$days_diff>=5,]
p <- ggplot(x,aes(factor(issue_severity),days_diff))
p + geom_violin() + geom_jitter(width = 0.2, alpha = 0.7) + 
  labs(title="Credit or prepaid card")
q <- ggplot(y,aes(factor(issue_severity),days_diff))
q + geom_violin() + geom_jitter(width = 0.2, alpha = 0.7) +
  labs(title="Credit reporting")
r <- ggplot(z,aes(factor(issue_severity),days_diff))
r + geom_violin() + geom_jitter(width = 0.2, alpha = 0.7) +
  labs(title="Debt collection")

```

# Model Building
# Baseline Ordinal regression

```{r}
drops <- c("product","zip_code","timely_response","consumer_disputed","lat","lng","issue","state")
card_df <-card_df[ , !(names(card_df) %in% drops)]
```

```{r}
colnames(card_df)
```

```{r}
sum(is.na(card_df))
```

```{r}
set.seed(42)
trainingRows <- sample(1:nrow(card_df), 0.75 * nrow(card_df))
trainingData <- card_df[trainingRows,]
testData <- card_df[-trainingRows, ]
```


```{r}
library(MASS)
#run the Ordinal regression model
polrMod = polr(issue_severity ~sub_product+tags+consumer_consent_provided+submitted_via+company_response_to_consumer+day+days_diff+narrative_count+covid_indicator, data=trainingData,Hess=TRUE)
#print the estimated values of the parameters for the ordinal regression model
summary(polrMod)
```

```{r}
predictedclass <- predict(polrMod,testData)
table(testData$issue_severity, predictedclass)
```

```{r}
mean(as.character(testData$issue_severity) != as.character(predictedclass))
```

```{r}
drops <- c("product","consumer_disputed","day","narrative_count","zip_code","issue","timely_response","tags","lat","lng","state")
debt_df <-debt_df[ , !(names(debt_df) %in% drops)]
```


```{r}
debt_df$issue_severity <- factor(debt_df$issue_severity,levels=c("low","medium","high"),ordered=TRUE)

set.seed(42)
trainingRows <- sample(1:nrow(debt_df), 0.7 * nrow(debt_df))
trainingData <- debt_df[trainingRows,]
testData <- debt_df[-trainingRows, ]

```


```{r}
library(MASS)
#run the Ordinal regression model
polrMod = polr(issue_severity ~sub_product+consumer_consent_provided+submitted_via+company_response_to_consumer+days_diff+covid_indicator, data=trainingData,Hess=TRUE)
#print the estimated values of the parameters for the ordinal regression model
summary(polrMod)
```

```{r}
predictedClass <- predict(polrMod,testData)
table(testData$issue_severity, predictedClass)
```


```{r}
mean(as.character(testData$issue_severity) != as.character(predictedClass))
```


```{r}
drops <- c("product","consumer_disputed","day","narrative_count","zip_code","issue","tags","days_diff","company_response_to_consumer","lat","lng","state")
credit_df <-credit_df[ , !(names(credit_df) %in% drops)]
```

# Unique values in columns
```{r}
unique(trainingData$company_response_to_consumer)
```

```{r}
set.seed(42)
trainingRows <- sample(1:nrow(credit_df), 0.7 * nrow(credit_df))
trainingData <- credit_df[trainingRows,]
testData <- credit_df[-trainingRows, ]
```

```{r}
library(MASS)
#run the Ordinal regression model
polrMod = polr(issue_severity ~sub_product+consumer_consent_provided+submitted_via+timely_response+covid_indicator, data=trainingData,Hess=TRUE)
#print the estimated values of the parameters for the ordinal regression model
summary(polrMod)
```

```{r}
predictedclass <- predict(polrMod,testData)
table(testData$issue_severity, predictedclass)
```


```{r}
mean(as.character(testData$issue_severity) != as.character(predictedclass))
```

